The Sub-Millimetre User Reduction Facility: an iterative map-maker for SCUBA-2

### 2010-11-24 smurf_paper.tex
add first figure and some description of how scuba-2 works

\documentclass[useAMS,usenatbib,usegraphicx,nofootinbib]{mn2e}

@@

% --- Some user defined macros ------------------------------------------------
\newcommand{\apj}{\rm ApJ}
\newcommand{\apjl}{\rm ApJL}
\newcommand{\apjs}{\rm ApJS}
\newcommand{\aaps}{\rm A$\&$AS}
\newcommand{\aap}{\rm A$\&$A}
\newcommand{\aapr}{\rm A$\&$AR}
\newcommand{\mnras}{\rm MNRAS}
\newcommand{\aj}{\rm Astron. J.}
\newcommand{\araa}{\rm ARAA}
\newcommand{\nat}{\rm Nature}
\newcommand{\pasj}{\rm PASJ}
\newcommand{\pasp}{\rm Publ. Astron. Soc. Pac.}
\newcommand{\ASP}{\rm ASP COnference Series}
\newcommand{\CASP}{\rm Comm. Astrophys. Space Phys.}
\newcommand{\astroph}{\rm astro-ph/}

\newcommand{\snr}{SNR}

\newcommand{\scuba}{SCUBA-2}

\def\lsim{\mathrel{\lower2.5pt\vbox{\lineskip=0pt\baselineskip=0pt
           \hbox{$<$}\hbox{$\sim$}}}}

@@

\author[Edward~L.~Chapin~et~al.]{
  \parbox[t]{\textwidth}{
    Edward~L.~Chapin$^{1}$\thanks{E-mail:~echapin@phas.ubc.ca},
    David~S.~Berry$^{2}$,
    Andrew~G.~Gibb$^{1}$,
    Tim~Jenness$^{2}$,
    Douglas~Scott$^{1}$
  }
  \\
  \\
  $^{1}$Dept. of Physics \& Astronomy, University of British Columbia,
  6224 Agricultural Road, Vancouver, B.C. V6T 1Z1, Canada\\
  $^{2}$JointAstronomy Centre, 660 N. A‘oh ̄ k ̄ Place, University Park, Hilo, Hawaii 96720, USA}

@@

\begin{abstract}
  We describe the Sub-Millimetre User Reduction Facility (SMURF), an
  iterative map-maker that was developed for the Submillimetre Common
  User Bolometer Array 2 (SCUBA-2).
\end{abstract}

\begin{keywords}
bla, bla, bla
\end{keywords}

@@

The Submillimetre Common User Bolometer Array 2
\citep[\scuba,][]{holland2006} is a new instrument for the the 15-m
James Clerk Maxwell Telescope (JCMT) on Mauna Kea, Hawai'i. The camera
can simultaneously image the sky in two broad bands centered over 450
and 850\,\micron. Ultimately, the instrument will also be equipped
with a fourier transform spectrometer (FTS), and a polarimeter
(POL). This paper describes the Submillimetre User Reduction Facility,
SMURF, a software package for reducing the imaging data, with an
emphasis on data taken during the SCUBA-2 Shared Risk Observing period
(S2SRO) which took place from February to April of 2010. The reduction
of FTS and POL data will be described at a later date once these
additional instruments have been commissioned.

@@

The most common source of emission at these wavelengths in the
Universe is the thermal emission from cool dust.

Styles of map-making: maximum likelihood
\citep[e.g.,][]{patanchon2008}; iterative -- see \citet{johnstone2000}
implementation of \citet{wright1996} to remove chop,
\citet{kovacs2008} for SHARC-2 etc.; de-correlation using PCA
analysis \citep[e.g.][]{laurent2005,scott2008}.

Since \scuba\ will ultimately have nearly 10,000 working bolometers,
considerably larger than any other existing ground, or space-spaced
bolometer cameras, SMURF has been designed to use a less accurate,
though faster and less memory-intensive iterative approach that
attempts to model and remove most of the correlated signal components,
and then regrid the residuals signals. This method is closest in
spirit to the Comprehensive Reduction Utility for SHARC-2
\citep[CRUSH,][]{kovacs2008}.

@@

A brief summary of the instrument hitting on topics that will be
discussed later: TES operation; base temperature and pixel heaters to
compensate for sky variations; measuring flatfields; sample rate; scan
speed to move astronomical sources of interest to higher frequencies
in the time series.

@@

This paper will be focussed on S2SRO: only one subarray at each
wavelength; fridge oscillations; working detectors.

@@

\subsection{Data cleaning}
We basically need to correct steps and apply the flatfield correction.

>>>>

\documentclass[useAMS,usenatbib,nofootinbib]{mn2e}

% this is needed to get pdftex to work properly!
\usepackage[pdftex]{graphicx}

@@

% --- Some user defined macros ------------------------------------------------

% journals
\newcommand{\apj}{\rm ApJ}
\newcommand{\apjl}{\rm ApJL}
\newcommand{\apjs}{\rm ApJS}
\newcommand{\aaps}{\rm A$\&$AS}
\newcommand{\aap}{\rm A$\&$A}
\newcommand{\aapr}{\rm A$\&$AR}
\newcommand{\mnras}{\rm MNRAS}
\newcommand{\aj}{\rm Astron. J.}
\newcommand{\araa}{\rm ARAA}
\newcommand{\nat}{\rm Nature}
\newcommand{\pasj}{\rm PASJ}
\newcommand{\pasp}{\rm Publ. Astron. Soc. Pac.}
\newcommand{\ASP}{\rm ASP COnference Series}
\newcommand{\CASP}{\rm Comm. Astrophys. Space Phys.}
\newcommand{\astroph}{\rm astro-ph/}

% common acronyms etc.
\newcommand{\snr}{SNR}
\newcommand{\scuba}{SCUBA-2}

% these are approximately less-than and greater-than symbols
\def\lsim{\mathrel{\lower2.5pt\vbox{\lineskip=0pt\baselineskip=0pt
          \hbox{$<$}\hbox{$\sim$}}}}

@@

\author[Edward~L.~Chapin~et~al.]{
  \parbox[t]{\textwidth}{
    Edward~L.~Chapin$^{1}$\thanks{E-mail:~echapin@phas.ubc.ca},
    David~S.~Berry$^{2}$,
    Andrew~G.~Gibb$^{1}$,
    Tim~Jenness$^{2}$,
    Douglas~Scott$^{1}$
  }
  \\
  \\
  $^{1}$Dept. of Physics \& Astronomy, University of British Columbia,
  6224 Agricultural Road, Vancouver, B.C. V6T 1Z1, Canada\\
  $^{2}$JointAstronomy Centre, 660 N. A`oh\={o}k\={u} Place, University
  Park, Hilo, Hawaii 96720, USA}

@@

\begin{abstract}
  We describe the properties of data from the Submillimetre Common
  User Bolometer Array 2 (SCUBA-2), and the production of maps using
  the Sub-Millimetre User Reduction Facility (SMURF).
\end{abstract}

\begin{keywords}
chicken, chicken: chicken, chicken: chicken -- chicken
\end{keywords}

@@

The Submillimetre Common User Bolometer Array 2
\citep[\scuba,][]{holland2006} is a new instrument for the the 15-m
James Clerk Maxwell Telescope (JCMT) on Mauna Kea, Hawai'i. The camera
can simultaneously image the sky in two broad bands centered over 450
and 850\,\micron. Ultimately, the instrument will also be equipped
with a Fourier transform spectrometer (FTS), and a polarimeter
(POL). This paper describes the Submillimetre User Reduction Facility,
SMURF, a software package for reducing the imaging data, with an
emphasis on data taken during the SCUBA-2 Shared Risk Observing period
(S2SRO) which took place from February to April of 2010. The reduction
of FTS and POL data will be described at a later date once these
additional instruments have been commissioned.

@@

Styles of map-making: maximum likelihood
\citep[e.g.,][]{patanchon2008}; iterative -- see \citet{johnstone2000}
implementation of \citet{wright1996} to remove chop,
\citet{kovacs2008} for SHARC-2 etc.; de-correlation using PCA analysis
\citep[e.g.][]{laurent2005,scott2008,aguirre2010}.

Since \scuba\ will ultimately have nearly 10,000 working bolometers,
considerably larger than any other existing ground, or space-spaced
bolometer cameras, SMURF has been designed to use a less accurate,
though faster and less memory-intensive iterative approach that
attempts to model and remove most of the correlated signal components,
and then regrid the residuals signals. This method is similar in
spirit to the Comprehensive Reduction Utility for SHARC-2
\citep[CRUSH,][]{kovacs2008}, and the pipeline developed for the
Bolocam Galactic Plane Survey \citep{aguirre2010}.

@@

Before describing the data-reduction algorithm in detail, we summarize
the basic strategy used to make measurements with SCUBA-2.

The first generation of submm bolometer cameras used NTD germanium
thermistors (e.g., SCUBA, MAMBO, BOLOCAM, AzTEC, LABOCA...).
Generally speaking, these devices are operated using a current bias;
changing thermal conditions alter their impedence, resulting in a
changing Voltage that can be measured. The bolometers are cooled so
that they are {\em background limited} -- the dominant noise term is
the optical load from sources outside of the cryostat (emissivity from
telescope, atmospheric fluctuations, and astronomical sources).

In contrast, SCUBA-2 belongs to a new generation of cameras that uses
superconducting transition-edge sensors (TESs, like ACT, SPT,
MUSTANG). These devices are held at the critical temperature, $T_C$,
below which the material becomes a superconductor, and above which it
becomes a normal conductor. Within this transition, tiny temperature
fluctuations can produce huge changes in impedence. The devices are
operated with a voltage bias, so that thermal fluctuations produce
changing currents, which in turn generate changing local magnetic
fields. Chains of superconducting quantum interference devices
(SQUIDs) are used to amplify these signals, ultimately providing a
larger changing current that may be digitized. Like the NTD
thermistor-based cameras, SCUBA-2 is cooled in order to be background
limited. However, a superior over-all sensitivity on a per-bolometer
basis is achieved through its improved responsivity (similar to an
improved quantum efficiency for a CCD).

The load on the detectors can vary substatially, both from thermal
sources within and outside the SCUBA-2 cryostat. In order to operate
the camera across a broad dynamic range of conditions, it is therefore
necessary to thermally control the focal plane. This is achieved by
providing a fridge {\em base temperature}, $T_B$ that is well-below
$T_C$, and then using heaters that are connected to each bolometer to
place the array in to transition.

The focal plane is divided into four quadrants, each filled with 32
column $\times$ 40 row sub-arrays labelled s4a--s4d at 450\,\micron,
and s8a--s8d at 850\,\micron. To avoid independent bias and readout
wires to each bolometer, each array is multiplexed by column. There is
an additional row of dark SQUIDs with no TESs or thermal absorber used
to monitor non-thermal noise sources.

The camera has been designed so that the absolute response of each
detector can be established by ramping currents through the individual
pixel heaters to produce Joule heating power, $P_\mathrm{J} =
\epsilon I^2R$, where $I$ is the known current, $R$ is the heater
resistance, and $\epsilon$ is the efficiency with which power is
transferred to the bolometer. In order to calibrate the total factor
$\epsilon R$, SCUBA-2 posesses an internal flatfield source.

For further detail see...

@@

This paper will be focussed on S2SRO: only one subarray at each
wavelength; fridge oscillations; working detectors (and variation from
setup-to-setup).

Show some time series.

Show power spectra.

Show cross-spectra to emphasize how correlated things are?

\begin{figure}
\centering
\includegraphics[width=0.99\linewidth]{bolos_point_mix.pdf}
\caption{A comparison between single bolometer time series in each
  band (top two plots), with the mixing chamber temperature and
  azimuth/elevation pointing offsets. There is a strong correlation
  between the bolometers and the roughly $\sim$30\,s oscillation in
  the fridge, but not with the telescope motion.}
\label{fig:bolos_mix}
\end{figure}

@@

We express the signal observed by the $i$th bolometer as a function of time,
\begin{equation}
B_i(t) = G_i[E_i(t) A_i(t) + N_i(t)]
\end{equation}
where $A(t)$ is the time-varying signal produced by scanning the
telescope across astronomical sources, $E(t)$ is the time-varying
extinction, which is a function of the telescope elevation and
atmospheric conditions, and $N_i(t)$ represents sources of noise. The
two terms in square brackets, as written, have units of power
delivered to the detectors (pW). The scale factor $G_i$ converts this
effective power to the digitized units recorded on disk (DAC).

In practice, $G_i$ is further factored into two components: $G_i = R_i
D$. The responsivity, $R_i$, gives the change in current as a function
of power fluctuations (A\,pW$^{-1}$). There is then a final (and
constant across all bolometers) conversion factor from current to DAC
units, $D$ (DAC/A).

\subsection{Data cleaning}

\subsubsection{Step correction}

\subsubsection{Flatfields}
\label{sec:flatfields}

\begin{itemize}

\item Heater ramps

\item common-mode signal amplitude

\item astronomical sources

\item flat-field source

\end{itemize}
